### 2018-08-06 paper.tex
Describing the first agent (almost done)

\documentclass[runningheads]{llncs}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{color}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{wasysym}
\usepackage[lighttt]{lmodern}
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\newcommand\Dist{\mathit{Dist}}
\newcommand\ctrl{\mathit{ctrl}}
\newcommand\prnt{\mathit{prnt}}
\newcommand\link{\mathit{link}}
\newcommand\points{\mathit{points}}
\newcommand\ports{\mathit{ports}}
\newcommand\ar{\mathit{ar}}
\newcommand\id{\mathsf{id}}
\newcommand\Id{\mathsf{Id}}

\providecommand\longrightarrowRHD{\relbar\joinrel\relbar\joinrel\mathrel\RHD}
\providecommand\longrightarrowrhd{\relbar\joinrel\relbar\joinrel\mathrel\rhd}

\makeatletter
\providecommand*\xrightarrowRHD[2][]{\ext@arrow 0055{\arrowfill@\relbar\relbar\longrightarrowRHD}{#1}{#2}}
\providecommand*\xrightarrowrhd[2][]{\ext@arrow 0055{\arrowfill@\relbar\relbar\longrightarrowrhd}{#1}{#2}}
\makeatother

\lstdefinelanguage{Big}{
  basicstyle=\ttfamily,
  sensitive=true,
  morekeywords={fun, brs, end, sbrs, pbrs, nbrs, begin, init, atomic, preds,
    rules, action, big, ctrl, float, int, react},
  comment=[l]{\#}
}

\ttfamily
\DeclareFontShape{OT1}{lmtt}{m}{it}
     {<->sub*lmtt/m/sl}{}

\begin{document}
\title{Nondeterministic Bigraphical Reactive Systems for Markov Decision
  Processes\thanks{Supported by organization x.}}
\titlerunning{Nondeterministic BRSs for MDPs}
\author{Paulius Dilkas \and Michele Sevegnani}
\authorrunning{P. Dilkas \& M. Sevegnani}
\institute{University of Glasgow, Glasgow, UK}
\maketitle

\begin{abstract}
The abstract should briefly summarize the contents of the paper in
150--250 words.

\keywords{Bigraphs \and Probabilistic semantics \and Markov decision process.}
\end{abstract}

\section{Introduction}

\subsection{Markov Decision Processes}

\begin{definition}[Markov decision process] \label{mdp}
  For any finite set $X$, let $\Dist(X)$ denote the set of discrete probability
  distributions over $X$. A \emph{Markov Decision Process} is a tuple $ (S,
  \overline{s}, A, P, L)$, where: $S$ is a finite set of states and
  $\overline{s} \in S$ is the initial state; $A$ is a finite set of
  \emph{actions}; $P : S \times A \to \Dist(S)$ is a (partial) probabilistic
  transition function, mapping state-action pairs to probability distributions
  over $S$; $L : S \to 2^P$ is a labelling with atomic propositions.
\end{definition}

\begin{definition}[rewards]
  A \emph{reward structure} for an MDP $(S, \overline{s}, A, P, L)$ is a pair
  $(\rho, \iota)$, where $\rho : S \to \mathbb{R}$ is the \emph{state reward
    function}, and $\iota : S \times A \to \mathbb{R}$ is the \emph{transition
    reward function}.
\end{definition}

\subsection{Bigraphs}

From \cite{dblp:journals/tcs/sevegnanic15}.
From PhD thesis \cite{DBLP:phd/ethos/Sevegnani12}

\begin{figure}
  \includegraphics{models/example_bigraph/foo_lhs.pdf}
  \caption{An example bigraph.}
  \label{example_bigraph}
\end{figure}

Consider the graphical representation of a bigraph in Fig.
\ref{example_bigraph}. The two white dashed rectangles represent \emph{regions},
i.e. parents of any otherwise parentless nodes. Grey dashed rectangles are
called \emph{sites} and encode parts of the model that have been abstracted
away. Nodes, sites, and regions are also known as \emph{places}. Two places with
the same parent, or two points with the same link are called \emph{siblings}.

\begin{definition}[concrete place graph with sharing]
  A \emph{concrete place graph with sharing}
  \[ F = (V_F, \ctrl_F, \prnt_F) : m \to n \]
  is a triple having an inner interface $m$ and an outer interface $n$. These
  index the sites and regions of the place graph, respectively. $F$ has a finite
  set $V_F \subset \mathcal{V}$ of nodes, a control map $\ctrl_F : V_F \to
  \mathcal{K}$, and a \emph{parent relation}
  \[ \prnt_F \subseteq (m \uplus V_F) \times (V_F \uplus n) \]
  that is acyclic, i.e., $(v, v) \not\in \prnt_F^+$ for any $v \in V_F$.
\end{definition}

\begin{definition}[composition for place graphs with sharing]
  If $F : k \to m$ and $G : m \to n$ are two concrete place graphs with sharing
  with $V_F \cap V_G = \emptyset$, their composite
  \[ G \circ F = (V, \ctrl, \prnt) : k \to n \]
  has nodes $V = V_F \uplus V_G$ and control map $\ctrl = \ctrl_F \uplus
  \ctrl_G$. Its parent relation $\prnt \subseteq (k \uplus V) \times (V \uplus
  n)$ is given by:
  \[ \prnt \coloneqq \prnt_G^\lhd \uplus \prnt_\circ \uplus \prnt_F^\rhd, \]
  where
  \begin{align*}
    \prnt_F^\rhd &= \prnt_F \rhd V_F, \\
    \prnt_G^\lhd &= V_G \lhd \prnt_G, \\
    \prnt_\circ &= (m \lhd \prnt_G) \circ (\prnt_F \rhd m).
  \end{align*}
\end{definition}

\begin{definition}[tensor product for place graphs]
  If $G_0 : m_0 \to n_0$ and $G_1 : m_1 \to n_1$ are two concrete place graphs
  with sharing with $V_F \cap V_G = \emptyset$, their tensor product
  \[ G_0 \otimes G_1 = (V, \ctrl, \prnt) : m_0 + m_1 \to n_0 + n_1 \]
  has nodes $V = V_{G_0} \uplus V_{G_1}$ and control map $\ctrl \coloneqq
  \ctrl_{G_0} \uplus \ctrl_{G_1}$. Its parent relation $\prnt \subseteq [(m_0 +
  m_1) \uplus V] \times [V \uplus (n_0 + n_1)]$ is defined as
  \[ \prnt_{G_0} \uplus \prnt_{G_1}^{(m_0, n_0)}, \]
  where
  \begin{align*}
    \prnt_{G_1}^{(m_0, n_0)} = &\{ (v, w) \mid (v, w) \in \prnt_{G_1} \text{ and } v, w \in V_{G_1} \} \\
    \uplus &\{ (m_0 + i, w) \mid (i, w) \in \prnt_{G_1}, w \in V_{G_1} \text{ and } i \in m_1 \} \\
    \uplus &\{ (v, n_0 + i) \mid (v, i) \in \prnt_{G_1}, v \in V_{G_1} \text{ and } i \in n_1 \} \\
    \uplus &\{ (m_0 + i, n_0 + j) \mid (i, j) \in \prnt_{G_1}, i \in m_1, \text{ and } j \in n_1 \}.
  \end{align*}
\end{definition}

\begin{definition}[concrete link graph]
  A \emph{concrete link graph}
  \[ F = (V_F, E_F, \ctrl_F, \link_F) : X \to Y \]
  is a quadruple having an inner face $X$ and an outer face $Y$, both finite
  subsets of $\mathcal{X}$, called respectivelly the inner and outer names of
  the link graph. $F$ has finite sets $V_F \subset \mathcal{V}$ of nodes and
  $E_F \subset \mathcal{E}$ of edges, a control map $\ctrl_F : V_F \to
  \mathcal{K}$ and a \emph{link map}
  \[ \link_F : X \uplus P_F \to E_F \uplus Y, \]
  where $P_F \coloneqq \{ (v, i) \mid i \in \ar(\ctrl_F(v)) \}$ is the set of
  ports of $F$. Thus $(v, i)$ is the $i$th port of node $v$. We shall call $X
  \uplus P_F$ the \emph{points} of $F$, and $E_F \uplus Y$ its \emph{links}.
\end{definition}

The sets of points and the set of ports of a link $l$ are defined by
\[ \points_F(l) \coloneqq \{ p \mid \link_F(p) = l \}, \quad \ports_F(l)
  \coloneqq \points_F(l) \setminus X. \]
An edge is \emph{idle} if it has no points. Identities over name sets are
defined by $\id_X = (\emptyset, \emptyset, \emptyset, \Id_X) : X \to X$.

\begin{definition}[concrete bigraph with sharing]
  A concrete bigraph
  \[ F = (V_F, E_F, \ctrl_F, \prnt_F, \link_F) : \langle k, X \rangle \to
    \langle m, Y \rangle \]
  consists of a concrete place graph with sharing $F^P = (V_F, \ctrl_F, \prnt_F)
  : k \to m$ and a concrete link graph $F^L = (V_F, E_F, \ctrl_F, \link_F) : X
  \to Y$. We write the concrete bigraph with sharing as $F = (F^P, F^L)$.
\end{definition}

Given a set of reaction rules, we refer to the configurations that a system may
adopt as \emph{states}. A bigraph with inner face $\epsilon$ is called
\emph{ground}.

\begin{definition}[solid bigraph]
  A bigraph is \emph{solid} if these conditions hold:
  \begin{enumerate}
  \item no regions or outer names are idle (a region is called \emph{idle}
    if it is empty);
  \item no two sites or inner names are siblings;
  \item the parent of every site is a node;
  \item no outer name is linked to an inner name.
  \end{enumerate}
\end{definition}

\begin{definition}[reaction rule] \label{reaction_rule}
  A \emph{reaction rule} is a pair
  \[ \mathsf{R} = (R : m \to J, R' : m \to J), \]
  sometimes written as $R \longrightarrowRHD R'$, where $R$ is the \emph{redex}
  and $R'$ the \emph{reactum}, and $R$ is solid. The rule generates all the
  \emph{ground reaction rules} $(r, r')$, where $r = (R \otimes \id_Y) \circ d$
  and $r' = (R' \otimes \id_Y) \circ d$ for some discrete ground
  \emph{parameter} $d : \epsilon \to \langle m, Y \rangle$. The \emph{reaction
    relation} $\longrightarrowrhd_{\mathsf{R}}$ over ground bigraphs is defined
  by
  \[ g \longrightarrowrhd_{\mathsf{R}} g' \text{ iff } g = D \circ r \text{ and
    } g' = D \circ r' \]
  for some bigraph $D$ and some ground reaction rule $(r, r')$ generated from
  $\mathsf{R}$.
\end{definition}

\begin{definition}[bigraphical reactive system (BRS)]
  A \emph{bigraphical reactive system} consists of a pair $(\mathcal{B},
  \mathcal{R})$, where $\mathcal{B}$ is a set of ground bigraphs and
  $\mathcal{R}$ is a set of reaction rules defined over $\mathcal{B}$. It has a
  reaction relation
  \[ \longrightarrowrhd_{\mathcal{R}} \coloneqq \bigcup_{\mathsf{R} \in
      \mathcal{R}} \longrightarrowrhd_{\mathsf{R}}, \]
  which will be written $\longrightarrowrhd$ when $\mathcal{R}$ is understood.
\end{definition}

\section{Extensions to Bigraphs}

\subsection{Probabilistic Bigraphs}

\begin{definition}[probabilistic reaction rule]
  A \emph{probabilistic reaction rule} $\mathsf{R}$ is a triple $(R, R', p)$,
  sometimes written $R \xrightarrowRHD{p} R'$, where $(R, R')$ is a reaction
  rule and $p \in (0, 1]$ is a probability. Similarly to
  Definition~\ref{reaction_rule}, it generates a set of ground reaction rules of
  the form $(r, r', p)$.
\end{definition}

\begin{definition}[probabilistic bigraphical reactive system (PBRS)]
  A \emph{probabilistic bigraphical reactive system} consists of a pair
  $(\mathcal{B}, \mathcal{R})$, where $\mathcal{B}$ is a set of ground bigraphs
  and $\mathcal{R}$ is a set of probabilistic reaction rules defined over
  $\mathcal{B}$.

  Let $g, g'$ be ground bigraphs, and $\{ (r_i, r'_i, p_i) \}_{i=1}^n$ a set of
  ground probabilistic reaction rules, where for each $r_i$, there exists a
  bigraph $D_i$ such that $g = D_ir_i$. Let $S = \{ (r_i, r'_i, p_i) \mid g' =
  D_ir'_i \}$ (for the same $D_i$), and
  \[ s = \sum_{i=1}^n p_i. \]
  Then the reaction relation is defined as
  \[ g \xrightarrowrhd{p}_{\mathcal{R}} g' \text{ iff } S \ne \emptyset, \]
  where
  \[ p = \frac{1}{s}\sum_{(r, r', p') \in S} p'. \]
\end{definition}

\subsection{Nondeterministic Bigraphs}

\begin{definition}[nondeterministic reaction rule]
  Let $A$ be a set of actions. A \emph{nondeterministic reaction rule}
  $\mathsf{R}$ is a tuple $(R, R', a, p)$, where $(R, R', p)$ is a probabilistic
  reaction rule, and $a \in A$ is an action. We also define a \emph{reaction
    reward function} $r : A \to \mathbb{R}$ that assigns a reward/cost to each
  action.
\end{definition}

\begin{definition}[nondeterministic bigraphical reactive system (NBRS)]
  A \emph{nondeterministic bigraphical reactive system} consists of a pair
  $(\mathcal{B}, \mathcal{R})$, where $\mathcal{B}$ is a set of ground bigraphs
  and $\mathcal{R}$ is a set of nondeterministic reaction rules defined over
  $\mathcal{B}$.

  Let $g, g'$ be ground bigraphs, $a \in A$ an action, and $\{ (r_i, r'_i, a,
  p_i) \}_{i=1}^n$ a set of ground nondeterministic reaction rules with action
  $a$, where for each $r_i$, there exists a bigraph $D_i$ such that $g =
  D_ir_i$. Let $S = \{ (r_i, r'_i, a, p_i) \mid g' = D_ir'_i \}$ (for the same
  $D_i$), and
  \[ s = \sum_{i=1}^n p_i. \]
  Then the reaction relation for action $a$ is defined as
  \[ g \xrightarrowrhd[r(a)]{p}_{a} g' \text{ iff } S \ne \emptyset, \]
  where
  \[ p = \frac{1}{s}\sum_{(r, r', a, p') \in S} p'. \]
\end{definition}

BigraphER \cite{DBLP:conf/cav/SevegnaniC16} supports defining predicates, i.e.,
bigraphs that are compared to every encountered state. We can associate a reward
with each predicate, allowing us to assign rewards to states in a flexible and
semantically meaningful way: the reward of a state is simply the sum of the
rewards of all matching predicates (and $0$ in case there are none).

\begin{proposition}
  Any MDP can be expressed as an NBRS.
\end{proposition}
\begin{proof}
  Obvious.
\end{proof}

\section{Implementation}

\begin{example} \label{example}
  Consider an MDP $(S, \overline{s}, A, P, L)$, where $S = \{ s_0, s_1, s_2, s_3
  \}$, $\overline{s} = s_0$, $A = \{ a, b, c \}$, and $P, L$ defined as follows:
  \begin{equation*}
    \begin{split}
      P(s_0, a) &= [s_1 \mapsto 1], \\
      P(s_1, b) &= [s_0 \mapsto 0.7, s_1 \mapsto 0.3], \\
      P(s_1, c) &= [s_2 \mapsto 0.5, s_3 \mapsto 0.5], \\
      P(s_2, a) &= [s_2 \mapsto 1], \\
      P(s_3, a) &= [s_3 \mapsto 1],
    \end{split}
    \qquad
    \begin{split}
      L(s_0) &= \{ initial \}, \\
      L(s_1) &= \emptyset, \\
      L(s_2) &= \{ heads \}, \\
      L(s_3) &= \{ tails \}.
    \end{split}
  \end{equation*}
  Furthermore, equip it with a reward structure $(\rho, \iota)$, where
  $\rho(s_2) = 3$, $\iota(s_1, b) = 1$, and both functions are zero everywhere
  else.
\end{example}

\begin{figure}
  \begin{tabular}{p{0.5\textwidth}p{0.5\textwidth}}
    \begin{minipage}{0.5\textwidth}
      \vspace{13mm}
      \lstinputlisting[caption=BigraphER code., captionpos=b, language=Big, label=example_listing]{models/example.big}
    \end{minipage}
    &
      \begin{minipage}{0.5\textwidth}
        \centering
        \includegraphics{models/example_ts.pdf}
        \caption{The full transition system.}
        \label{example_ts}
      \end{minipage}
  \end{tabular}
\end{figure}

The MDP can be represented as an NBRS with BigraphER code in Listing
\ref{example_listing}. Furthermore, the BigraphER's visualisation of the
(automatically generated) transition system is in Fig. \ref{example_ts}.

More specifically, reaction rule probabilities are represented as floating-point
numbers inside the arrows (e.g. \texttt{-[0.7]->}), each action encompasses its
reaction rules with \texttt{\textbf{action} actionName} and
\texttt{\textbf{end}}. Rewards can be added to actions simply by inserting an
integer enclosed in squared brackets after the action name (e.g.
\texttt{\textbf{action} b[1]}). Lastly, predicate rewards have the same format,
but are listed on the \texttt{\textbf{preds}} line of the \texttt{\textbf{begin
    nbrs}}---\texttt{\textbf{end}} block (e.g. \texttt{\textbf{preds} =
  \{heads[3]\};}).

Visually, each state is represented by a white ellipse (the starting state is
highlighted with a thicker border), and each action (per state) is a grey
ellipse. Text inside state ellipses lists all predicates that are satisfied by
that state. Edges from actions to states are labelled with both the probability
and the name of the reaction rule. Transitions from either a fully generated
transition system or a simulation, labels from predicates, and state/transition
rewards can be exported to the corresponding PRISM plain text formats.

% TODO: introduce this
\begin{figure}
  \includegraphics{images/interface.jpg}
  \caption{The BigraphER Jupyter interface with syntax highlighting.}
  \label{interface}
\end{figure}

\section{A Case Study in Autonomous Agents}

Akin to \cite{dblp:conf/nfm/giaquintahimn18}.

\subsection{\#1}

Grid layout (but could be any nodes-and-edges layout, as long as it's clear how
to go back home). Tracking visited/unvisited states. Each unvisited state has a
probability $p$ of containing an object. The agent returns (and stays) home
after collecting a set number of objects (easily customisable). Starting at the
bottom left cell.

\lstinputlisting[caption=Controls., captionpos=b, language=Big,
label=agent1_controls, firstline=3, lastline=17]{models/agent1.big}

We begin by introducing controls: the types of nodes, what they contain, and
what they are linked to. The full list of controls is in Listing
\ref{agent1_controls}. \texttt{Cell} represents a discrete location. It is
linked to either \texttt{Visited} or \texttt{Unvisited} (initially, all except
one cells are unvisited). \texttt{Cell}s always contain
\texttt{Directions}, and one \texttt{Cell} contains an \texttt{Agent}. The
\texttt{Agent} may contain an \texttt{Object}, which may contain another
\texttt{Object}, etc. \texttt{Directions} always contain four nodes:
\texttt{North}, \texttt{East}, \texttt{West}, \texttt{South}. A \texttt{West}
node can be linked either to one of the two \texttt{Void}s or to an
\texttt{East} node of a neighbouring \texttt{Cell} (and similarly with other
directions). \texttt{Void}s represent the edges of the map. Using two
\texttt{Void} controls prevents us from unintentionally connecting the opposite
edges of the map. \texttt{Void1} corresponds to top and left edges, while
\texttt{Void2} is for bottom and right.

\lstinputlisting[caption=Predicates and the initial state., captionpos=b,
language=Big, label=agent1_code, firstline=19,
lastline=50]{models/agent1.big}

Next, we define two predicates: \texttt{home} and \texttt{goal} (see Listing
\ref{agent1_code}). The former states that the \texttt{Agent} is home
if it is in the southwest corner of the grid,
while the latter says that the agent should go home as soon as
it gets one object. A larger number of objects can be set by,
e.g., \texttt{Agent.Object.Object}. Additionally, Listing \ref{agent1_code}
defines the initial $2 \times 2$ grid, as described previously.

\lstinputlisting[caption=The action of going \texttt{north}, captionpos=b,
language=Big, label=agent1_north, firstline=108,
lastline=141]{models/agent1.big}

Lastly, we have five actions (four for the directions, and one for
going home). Each of the actions for directions carries three reaction
rules: one for going to an already visited cell, one for visiting a
cell for the first time, and one for visiting a cell for the
first time and finding an object. We will use the rules for going north
in Listing \ref{agent1_north} as an example. When going to an already visited
cell, the \texttt{Agent} simply traverses a link between the \texttt{North} node
of its previous \texttt{Cell} and the corresponding \texttt{South} node. As this
is the only reaction rule for this situation, its probability is $1$. When
going to an unvisited node, we equip an \texttt{Object} with probability $p$.
Regardless of whether that happens or not, the \texttt{Cell} changes its
status from \texttt{Unvisited} to \texttt{Visited}. % TODO: going home now

\section{Conclusion}

\bibliographystyle{splncs04}
\bibliography{bibliography}
\end{document}